---
title: "COMP3008 - Big Data Analytics"
author: "Workshop 7 - Introduction to VisualisationR"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This worksheet is designed to give you an introduction to creating visualisations in R. R is a programming language for statistical computing and data visualization. It has been adopted in the fields of data mining, bioinformatics, and data analysis.

Like Python, R uses libraries of functions to carry out tasks which can be imported as follows:

```{r}
library(readr) # Loads in the library readr
```

## Exercise 1 - Plotting data using ggplot2

A small company has recorded the following data on volume and costs of production for the last ten months:

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|llllllllll|}
\hline
Production (1,000 units)& 2& 4 &6 &6 &10& 8& 5& 7& 11& 12\\
\hline
Costs (£1,000s) &7 &11& 12& 19& 22& 20& 16& 13& 24& 20\\
\hline
\end{tabular}
\end{center}
\end{table}

Create a data frame containing variables production and costs, then plot these data. You will need to load in the library \texttt{ggplot2}.

You should end up with the following:

\textbf{Solution}

```{r,warning=FALSE}
production<-c(2, 4,6,6,10, 8, 5, 7, 11, 12)
costs<-c(7,11, 12, 19, 22, 20, 16, 13, 24, 20)

df<-as.data.frame(cbind(production,costs))

library(ggplot2)

ggplot(df,aes(x=production,y=costs))+geom_point()
```

Now plot production against time.To do this you will need to produce a sequence of numbers. Create a sequence from 1 to 10 and call it month. Then include month in your data frame so that it now contains production, costs and the newly created month.

\textbf{Solution}

```{r}
month<-c(1:10)
df<-cbind(df,month)

ggplot(df,aes(x=month,y=production))+geom_point()
```

The package \texttt{ggplot2} offers large flexibility of plotting options and control over the plot. Each function has a help page available which can be accessed like so:

```{r,eval=FALSE}
?mean()
```

Have look at the help file for \texttt{ggplot2} and customise your plot. On my plot below I have added a linear model and specified the colour of the line, made the points bigger and changed the colour, increased the size and changed the colour of the axis labels.

\textbf{Solution}

```{r,message=FALSE}

ggplot(df,aes(x=month,y=production))+geom_point(size=5,col="hotpink")+
  geom_smooth(method="lm",col="red")+
  theme(axis.title.x = element_text(size=16, colour = "blue"),                            
        axis.title.y = element_text(size=20, colour = "darkgreen"))
```


## Exercise 2 - Reading data into R

On the DLE is a file named \texttt{imbd\_top\_1000\_movies.csv}, use the function \texttt{read\_csv} to load the file into R. 

Once the data is loaded in, you can use the functions \texttt{head} and \texttt{tail} to view the first and last 6 rows of data.

\textbf{Solution}

```{r}
imdb_data<-read.csv("imdb_top_1000_movies.csv")

head(imdb_data)

tail(imdb_data)

```

## Exercise 3

First look at the structure of the data:

```{r}
str(imdb_data)
```

We see that runtime is currently being stored as a character string. We want to extract the numerical value from this.

```{r}

imdb_data$Runtime<-gsub(" min","",imdb_data$Runtime) ## Removes the character elements

imdb_data$Runtime<-as.numeric(imdb_data$Runtime) ## converts to numerical data.
imdb_data$Released_Year<-as.numeric(imdb_data$Released_Year)

```

Now we can perform the analysis on the data

```{r}
library(dplyr)

runtime_summary<-imdb_data %>% select(Runtime,Released_Year) %>% 
  group_by(Released_Year) %>% summarise(ave_runtime=mean(Runtime))

```

Produce a plot to show this:

```{r,warning=FALSE}

ggplot(runtime_summary,aes(x=Released_Year,y=ave_runtime))+geom_line()+
  labs(title = "Plot of the average rumtime for movies in the IMDB top 1000 movies database")+
  xlab("Year of Release")+ylab("Average runtime (mins)")+
  theme(axis.title.x = element_text(size=14),                            
        axis.title.y = element_text(size=14))

```