---
title: "COMP3008 - Big Data Analytics"
author: "Workshop 8 - Sentiment Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this worksheet we will be using R to perform sentiment analysis and produce different visualisations of the data. All the data required for this practical is available on the DLE. 


## Exercise 1 - Word Clouds

The first task is to create word clouds for the data contained in the two files, \texttt{Character\_A.txt} and \texttt{Character\_B.txt} using the following sentiments 

- Bing
- Afinn
- nrc
- Loughran

The sentiment are available in the \texttt{tidytext} and \texttt{textdata} packages and you can read the data in as follows:

```{r,eval=FALSE}
character_a<-read.delim("Character_A.txt", header = FALSE)

```

Make sure to include \texttt{header=FALSE} to ensure that none of the text ends up in the column name of the dataframe.

Is there a difference in the words contained in each of the lexicons? Is there a difference in the words each of the character uses?

Do you agree with how all the words have been categorised in each lexicon? 

\textbf{Solution}

```{r,warning=FALSE,message=FALSE}

## Load in the required libraries
library(tidyverse)
library(tidytext)
library(textdata)
library(wordcloud)
library(wordcloud2)

```

```{r,warning=FALSE}

## Load in the data 

character_a<-read.delim("Character_A.txt", header = FALSE)
character_b<-read.delim("Character_B.txt", header = FALSE)

## Tokenise the data (put one word into one row)

character_a<- character_a %>% select(V1) %>% ## column that contains the text data
  unnest_tokens(word, V1)

character_b<- character_b %>% select(V1) %>% ## column that contains the text data
  unnest_tokens(word, V1)

## Ckeck the data

head(character_a)
head(character_b)

```


```{r,warning=FALSE,message=FALSE}

## Join the tokenised data with the sentiments and produce the word clouds

## First for character a

character_a %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red"),
                   max.words = 100)


character_a %>%
  inner_join(get_sentiments("afinn")) %>%
  count(word, value, sort = TRUE) %>%
  reshape2::acast(word ~ value, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen"),
                   max.words = 100)


character_a %>%
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen",
                              "darkblue","darkmagenta"),
                   max.words = 100)


character_a %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen",
                              "darkblue","darkmagenta"),
                   max.words = 100)

```

```{r,warning=FALSE,message=FALSE}

## Now for character b

character_b %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red"),
                   max.words = 100)


character_b %>%
  inner_join(get_sentiments("afinn")) %>%
  count(word, value, sort = TRUE) %>%
  reshape2::acast(word ~ value, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen"),
                   max.words = 100)


character_b %>%
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen",
                              "darkblue","darkmagenta"),
                   max.words = 100)


character_b %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("blue", "red","gold","green","purple","hotpink","seagreen",
                              "darkblue","darkmagenta"),
                   max.words = 100)

```


Is there a difference in the words contained in each of the lexicons? Is there a difference in the words each of the character uses?

\textcolor{blue}{Character a uses the words connected to crime such as addict, assault and crime where as character b frequently uses words to describe themselves/others such as threatening, angry and pain. The word like and right are both the most common positive word and perhaps we should think about removing these to better identify the differences between the characters. The characters also have a difference in the most frequent word in the Afinn lexicon which are at opposite ends of the sentiment spectrum.}

Do you agree with how all the words have been categorised in each lexicon? 

\textcolor{blue}{For both these characters the words appear to be classified correctly but if the text was known this would give extra insight.}


## Exercise 2 - Sentiment Distribution

For this exercise you will need to use the data in the \texttt{Music\_data.csv} file. This file contains the lyrics to each of the songs from an album. Using the Bing and Afinn lexicons calculate the total sentiment for each song and plot the distribution of the sentiment for the album using a histogram.

To create a value from the bing lexicon, you will need the following code:

```{r,eval=FALSE}

toal_sentiment <- cleaned_music_words %>%
  inner_join(get_sentiments("bing")) %>% ## get the sentiment for the words
  count(tracknumber, sentiment) %>% ## count the sentiments for each track
  spread(sentiment, n, fill = 0) %>% ## negative and positive sentiment in separate columns
  mutate(score = positive - negative) ## score = net sentiment (positive - negative)

```

**Hint**: you will need to create a new column to denote the track number.

\textbf{Solution}

```{r}
## Read in the data 

music_data<-read.csv("Music_data.csv")

## Again we need to tokenise the data but keep the track data information

cleaned_music_words <- music_data %>%
  select(Text) %>% 
  mutate(tracknumber = row_number()) %>% # create new variable denoting the track number
  unnest_tokens(word, Text)

head(cleaned_music_words)

```

```{r}
## Distribution of sentiment using the bing lexicon

bing_sentiment <- cleaned_music_words %>%
  inner_join(get_sentiments("bing")) %>% ## get the sentiment for the words
  count(tracknumber, sentiment) %>% ## count the sentiments for each track
  spread(sentiment, n, fill = 0) %>% ## negative and positive sentiment in separate columns
  mutate(score = positive - negative) ## score = net sentiment (positive - negative)


## We can also calculate the mean and add this to the plot

bing_sentiment_mean <- bing_sentiment %>% 
  summarize(mean_score = mean(score)) 

ggplot(bing_sentiment, 
       aes(x = score)) + # Sentiment score on x-axis
  geom_bar(fill = "lightgreen", colour = "darkgreen") + # geom_bar will do the tabulation for you :-)
  geom_vline(aes(xintercept = mean_score), data = bing_sentiment_mean) +
  # Add a vertical line at the mean score, calculated and stored in sentiment_means_nba above
  geom_text(aes(x = mean_score, 
                y = Inf, 
                label = signif(mean_score, 3)), # Show to three significant figures
            vjust = 1.3, 
            data = bing_sentiment_mean) + 
  # Add the mean as a number; vjust moves it down from the top of the plot
  scale_x_continuous(breaks = seq(-15,25,by=5),  # Specify a suitable integer range for the x-axis
                     minor_breaks = NULL) + # Show integers; set this to a suitably large range
  labs(title = paste("Sentiments of tracks give a mean of", signif(bing_sentiment_mean$mean_score, 3)),
       # Title that gives page name and mean sentiment score, to three significant figures
       x = "Sentiment Score (Bing)", 
       y = "Number of tracks")+
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size=12),
        axis.text = element_text(size=12))


```

```{r}
## Distribution of sentiment using the Afinn lexicon

afinn_word_counts <- cleaned_music_words %>%
  inner_join(get_sentiments("afinn")) %>%
  count(tracknumber,word, value, sort = TRUE) %>%
  mutate(word = reorder(word, n)) 


afinn_sentiment<-afinn_word_counts%>%group_by(tracknumber)%>%
  mutate(total=sum(value*n))

## Remove duplicate values

afinn_sentiment<-afinn_sentiment[!duplicated(afinn_sentiment$tracknumber),]


afinn_sentiment_mean <- as.data.frame(mean(afinn_sentiment$total))   

names(afinn_sentiment_mean)[1]<-"mean_score"

ggplot(afinn_sentiment, 
       aes(x = total)) + # Sentiment score on x-axis
  geom_bar(fill = "lightgreen", colour = "darkgreen") + # geom_bar will do the tabulation for you :-)
  geom_vline(aes(xintercept = mean_score), data = afinn_sentiment_mean) +
  # Add a vertical line at the mean score, calculated and stored in sentiment_means_nba above
  geom_text(aes(x = mean_score, 
                y = Inf, 
                label = signif(mean_score, 3)), # Show to three significant figures
            vjust = 1.3, 
            data = afinn_sentiment_mean) + 
  # Add the mean as a number; vjust moves it down from the top of the plot
  scale_x_continuous(breaks = seq(-25,55,by=5),  # Specify a suitable integer range for the x-axis
                     minor_breaks = NULL) + # Show integers; set this to a suitably large range
  labs(title = paste("Sentiments of tracks give a mean of", signif(afinn_sentiment_mean$mean_score, 3)),
       # Title that gives page name and mean sentiment score, to three significant figures
       x = "Sentiment Score (Afinn)", 
       y = "Number of tracks")+
  theme(plot.title = element_text(size = 14),
        axis.title = element_text(size=12),
        axis.text = element_text(size=12))


```

## Exercise 3 - Sentiment Over Time

For this exercise you will need to use the data in the \texttt{Music\_data.csv} file and the \texttt{worldcup2022\_tweets.csv} file (data source : <https://www.kaggle.com/datasets/konradb/qatar-world-cup-2022-tweets>).

Using these datasets plot the sentiment over time/track number as a line plot. For the Twitter data, you will need to calculate the total sentiment for each tweet and group together by the date created.

**Hint**: the timestamp will read in as a character string, you can convert it using the \texttt{as.POSIct()} function.


\textbf{Solution}


```{r}

## The music data first as it is already cleaned and in the correct format to plot

## Bing sentiment

ggplot(bing_sentiment,aes(x=tracknumber,y=score))+geom_line()+labs(x="Track number",
                                                                   y="Sentiment score (Bing)")

## Afinn sentiment

ggplot(afinn_sentiment,aes(x=tracknumber,y=total))+geom_line()+labs(x="Track number",
                                                                   y="Sentiment score (Afinn)")

```


```{r,warning=FALSE,message=FALSE}

## Read in the data

worldcupdata<-read.csv("worldcup2022_tweets.csv")


## Look at the structure to see haw the timestamp has been encoded

str(worldcupdata$date)

## The date variable has been stored as a character so we need to convert it

worldcupdata$date<-as.POSIXct(worldcupdata$date, format="%Y-%m-%d")

## Check the format is correct. I have just kept the year, month and day but you can keep the 
## time by altering the format part above

str(worldcupdata$date)

cleaned_tweets <- worldcupdata %>%
  select(date,text) %>% 
  mutate(tweetnumber = row_number()) %>% # create new variable denoting the track number
  unnest_tokens(word, text)


## Now get the sentiments as before

worldcup_bing<-cleaned_tweets %>%
  inner_join(get_sentiments("bing")) %>% ## get the sentiment for the words
  count(date, sentiment) %>% ## count the sentiments for each day
  spread(sentiment, n, fill = 0) %>% ## negative and positive sentiment in separate columns
  mutate(score = positive - negative) ## score = net sentiment (positive - negative)



ggplot(worldcup_bing,aes(x=date,y=score))+geom_line()+labs(x="Date",y="Sentiment score (Bing)")


worldcup_afinn <- cleaned_tweets %>%
  inner_join(get_sentiments("afinn")) %>%
  count(date,word, value, sort = TRUE) %>%
  mutate(word = reorder(word, n)) 

worldcup_afinn<-worldcup_afinn%>%group_by(date)%>%
  mutate(total=sum(value*n))


worldcup_afinn<-worldcup_afinn[!duplicated(worldcup_afinn$date),]


ggplot(worldcup_afinn,aes(x=date,y=total))+geom_line()+labs(x="Date",y="Sentiment score (Afinn)")

```