---
title: "COMP3008 - Big Data Analytics"
author: "Workshop 6 - Data Visualisation Part 1"
date: "2026-01-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This worksheet is designed to give you an introduction to creating visualisations in R. R is a programming language for statistical computing and data visualization. It has been adopted in the fields of data mining, bioinformatics, and data analysis.

Like Python, R uses libraries of functions to carry out tasks which can be imported as follows:

```{r, warning=FALSE}
library(readr) # Loads in the library readr
```

In this workshop we will be using the IMDB database (\texttt{imbd\_top\_1000\_movies.csv}) and only the plotting functions with R. We bring the data into your environment using the code below:

```{r}

IMDB_data<-read_csv("imdb_top_1000_movies.csv")

```


Once the data is loaded in, you can use the functions \texttt{head} and \texttt{tail} to view the first and last 6 rows of data to ensure it has been loaded in correctly.


## Exercise 1

As a statistical software, R can produce a number of different types of plots without the need for external libraries. 

The first exercise is to use the \texttt{plot()} function with just the IMDB Rating column.

You should end up with the following:

```{r,echo=FALSE}
plot(IMDB_data$IMDB_Rating)
```
Not the most intuitive plot.

## Exercise 2

We can specify which particular type of plot we want R to make by using the plot name. Recreate the plot below using the \texttt{barplot()} function and the IMDB Rating variable.

```{r,echo=FALSE}
barplot(IMDB_data$IMDB_Rating)
```



## Exercise 3

Basic plotting in R has a simple syntax, you can find a cheatsheet for basic plotting [here](https://publish.illinois.edu/johnrgallagher/files/2015/10/BaseGraphicsCheatsheet.pdf). 


Using the IMBD data and the base R plotting functions, create the most hideous, misleading plot you can. If you wish to submit your wonderful creation, and I will show case your wonderful creations.
